Eagle: multi-locus association mapping on a genome-wide scale made routine

### 2019-08-19 BIOINF-2019-0261-Revision/bioinformaticspaper.tex
still tring to get bioinformatics in the right latex format

%%\documentclass{nature}
\documentclass{article}
\usepackage{xr}
\bibliographystyle{natbib}
\renewcommand{\baselinestretch}{1.4} 
\usepackage{graphicx}
\usepackage{bm}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{authblk}
\usepackage{xcolor}

@@

%% external tex files
%%\externaldocument{figure1_runtimes}
%%\externaldocument{figure2_powerFDRmultiple}
%%\externaldocument{figure3_mouse}

\externaldocument{supmaterials}

@@

\begin{abstract} \noindent
\setstretch{1.6}{
{\bf Motivation:} We present Eagle, a new method for multi-locus association mapping. The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. Eagle's strengths are that it a. is considerably more powerful than single-locus association mapping b. does not suffer from multiple testing issues c. gives results that are immediately interpretable and d. has a computational footprint comparable to single-locus association mapping. \\
{\bf Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods. We also analyse data from a published mouse study. Eagle found over 100\% more validated findings than the state-of-the-art single-locus method.  \\
{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R. It is freely available via the CRAN website  at https://cran.r-project.org. Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
{\bf Contact:} andrew.george@csiro.au
}
\end{abstract}

@@

We compared Eagle, in terms of computational and statistical performance, against seven other association mapping methods. 
We chose methods that almost all had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs or packages. Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multi-locus models. Of the many ways of performing single-locus association mapping, we chose 
GEMMA and FaST-LMM  because of their popularity and computational speed. 
For multi-locus association mapping, we chose bigRR, glmnet, 
LMM-Lasso, MLMM, and r2VIM.  
Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
is given in Supplementary Table \ref{suptabsummary} (see Methods for further details).

@@

The power and false positive rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure \ref{supfigpowermultiple}.  We restricted our attention to these scenarios because not all 
multi-locus methods could cope with the size of data in the other scenarios. 
  Each plot contains 
single points and power curves. The single points are the power and false positive rates for Eagle and MLMM.
These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
the false positive rate as the significance threshold  is adjusted. 
The power and false positive rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
 Figure \ref{figpowersingle}.

@@

In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
well.  Of the multi-locus methods, Eagle had the highest power
while keeping its false positive rate low (Supplementary Figure \ref{supfigpowermultiple}). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
association with a trait while avoiding false associations (Figure \ref{figpowersingle}).

@@

Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM. Not all data generated under the six scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs/packages was the limiting factor (see Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multi-locus programs were memory intensive. Most required in 
excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
Of the multi-locus programs/packages, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from our largest scenario, 10000 x 1.5M.

@@

The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}, 
\textcolor{red}{
with the Manhattan plots of the 
single-locus analysis shown in Supplementary Figure \ref{supmanhattan}.} 
The mouse study recorded
measurements on 200 traits. Of these, in the original study, 45 were able to have their findings  corroborated by previously published work. We focused 
our analyses here on these same 45 traits. Overall,  FaST-LMM$^{Bonf}$,  FaST-LMM$^{perm}$, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 47, 68, 37, 67, 106, SNP-trait findings, respectively, across 39 traits. No associations were found by FaST-LMM and Eagle for the other six traits. 
Eagle$^{alt}$ and Eagle$^{optimal}$ also found SNP-trait associations not found in the original study. This is despite their analyses being based 
on the SNP genotype data and the original study being based on SNP dosage data. Eagle$^{alt}$ found two  and 
Eagle$^{optimal}$ found seven new findings (Supplementary Table \ref{suptabnew}).  These new findings all involved SNPs whose association had been confirmed for other related traits in the original 
study.

@@

%%\bibliographystyle{naturemag}
\bibliography{biblibrary}

>>>>

%%\documentclass{nature}
\documentclass{article}
\usepackage{xr}

\renewcommand{\baselinestretch}{1.4} 
\usepackage{graphicx}
\usepackage{bm}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{authblk}
\usepackage{xcolor}

@@

@@

\begin{abstract} \noindent
\setstretch{1.6}{
{\bf Motivation:} We present Eagle, a new method for multi-locus association mapping. The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. Eagle's strengths are that it a. is considerably more powerful than single-locus association mapping b. does not suffer from multiple testing issues c. gives results that are immediately interpretable and d. has a computational footprint comparable to single-locus association mapping. \\
{\bf Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods. We also analyse data from a published mouse study. Eagle found over 50\% more validated findings than the state-of-the-art single-locus method.  \\
{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R. It is freely available via the CRAN website  at https://cran.r-project.org. Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
{\bf Contact:} andrew.george@csiro.au
}
\end{abstract}

@@

We compared Eagle, in terms of computational and statistical performance, against seven other association mapping methods. 
We chose methods that almost all had been purpose built for genome-wide analysis, that could handle data from quantitative traits, and where the methods had been implemented in freely available computer programs or packages. Two of the methods are based on single-locus (or locus-by-locus) models and five are based on multi-locus models. Of the many ways of performing single-locus association mapping, we chose 
GEMMA and FaST-LMM  because of their popularity and computational speed. 
For multi-locus association mapping, we chose bigRR, glmnet, 
LMM-Lasso, MLMM, and r2VIM.  
Each takes a different approach to multi-locus association mapping. A summary of the key attributes of the different computer programs/packages 
is given in Supplementary Table 1 (see Methods for further details).

@@

The power and false positive rates of Eagle and the other multi-locus methods across the scenarios 150 x 5K, 350 x 500K, 1500 x 50K, and 2000 x 500K are shown in Supplementary Figure 1.  We restricted our attention to these scenarios because not all 
multi-locus methods could cope with the size of data in the other scenarios. 
  Each plot contains 
single points and power curves. The single points are the power and false positive rates for Eagle and MLMM.
These two methods treat association mapping as a model selection problem. Their are no significance thresholds to be set. 
The power curves are for those methods that treat association mapping as a variable selection problem. Here, the 
significance of the findings are assessed against a significance threshold. The power curves in the plot show how power changes with 
the false positive rate as the significance threshold  is adjusted. 
The power and false positive rate of Eagle and the two single-locus methods, GEMMA and FaST-LMM,  are shown in
 Figure \ref{figpowersingle}.

@@

In answer to the question of how well Eagle finds true SNP-trait associations and avoids false SNP-trait associations, it does extremely 
well.  Of the multi-locus methods, Eagle had the highest power
while keeping its false positive rate low (Supplementary Figure 1). MLMM also performed well. However, it was when Eagle was compared against single-locus methods 
that the difference in power was most noticeable.  Eagle had much higher power than single-locus methods for finding SNP in true 
association with a trait while avoiding false associations (Figure \ref{figpowersingle}).

@@

Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM. Not all data generated under the six scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs/packages was the limiting factor (see Supplementary Figure 2).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multi-locus programs were memory intensive. Most required in 
excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
Of the multi-locus programs/packages, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from our largest scenario, 10000 x 1.5M.

@@

The genome wide results from the analyses of the mouse data are shown in Figure  \ref{figmouse}, 
\textcolor{red}{
with the Manhattan plots of the 
single-locus analysis shown in Supplementary Figure 3.} 
The mouse study recorded
measurements on 200 traits. Of these, in the original study, 45 were able to have their findings  corroborated by previously published work. We focused 
our analyses here on these same 45 traits. Overall,  FaST-LMM$^{Bonf}$,  FaST-LMM$^{perm}$, Eagle$^{default}$, Eagle$^{alt}$, and Eagle$^{optimal}$ found 47, 68, 37, 67, 106, SNP-trait findings, respectively, across 39 traits. No associations were found by FaST-LMM and Eagle for the other six traits. 
Eagle$^{alt}$ and Eagle$^{optimal}$ also found SNP-trait associations not found in the original study. This is despite their analyses being based 
on the SNP genotype data and the original study being based on SNP dosage data. Eagle$^{alt}$ found two  and 
Eagle$^{optimal}$ found seven new findings (Supplementary Table 2).  These new findings all involved SNPs whose association had been confirmed for other related traits in the original 
study.

@@

\bibliographystyle{natbib}
%%\bibliographystyle{naturemag}
\bibliography{biblibrary}

### 2019-08-19 BIOINF-2019-0261-RevisionV2/bioinformaticspaper.tex
still tring to get bioinformatics in the right latex format

%%\documentclass{nature}
\documentclass{article}
\usepackage{xr}

\renewcommand{\baselinestretch}{1.4} 
\usepackage{graphicx}
\usepackage{bm}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{authblk}
\usepackage{xcolor}

\linenumbers

\begin{document}

\title{Eagle: multi-locus association mapping on a genome-wide scale made routine}
\author[1]{Andrew W. George}
\author[2]{Arunas Verbyla}
\author[3]{Joshua Bowden}
\affil[1]{Data61, CSIRO, Brisbane, 4102, Australia.}
\affil[2]{Data61, CSIRO, Atherton, 4883, Australia.}
\affil[3]{IM \&T, CSIRO, Brisbane, 4067, Australia}
\date{}

@@

\begin{abstract} \noindent
\setstretch{1.6}{
{\bf Motivation:} We present Eagle, a new method for multi-locus association mapping. The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. Eagle's strengths are that it a. is considerably more powerful than single-locus association mapping b. does not suffer from multiple testing issues c. gives results that are immediately interpretable and d. has a computational footprint comparable to single-locus association mapping. \\
{\bf Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods. We also analyse data from a published mouse study. Eagle found over 50\% more validated findings than the state-of-the-art single-locus method.  \\
{\bf Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R. It is freely available via the CRAN website  at https://cran.r-project.org. Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
{\bf Contact:} andrew.george@csiro.au
}
\end{abstract}

\section{Introduction}
Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
 followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
linear regression. These designs though had a tendency to yield spurious findings if there was unrecognised population stratification 
\citep{cardon2003population}. This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \citep{spielman1996tdt}. Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \citep{price2010new}. This has meant that data can be collected from general populations, even if these populations are highly structured. Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine  \citep{yu2006unified,zhao2007arabidopsis}.

@@

\begin{figure}
\caption{
\setstretch{1.6}{
Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
FaST-LMM was run where all the SNP data are used 
to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
}
\vspace{-3cm}
\label{figpowersingle}
%%\begin{center}
\includegraphics[width=10cm, height=15cm]{power2main.eps}
%%\end{center}

@@

\begin{figure}
\caption{
\setstretch{1.6}{
Median run times, in minutes, for the analysis of simulation study data from the six scenarios. Eagle is compared against five other multi-locus programs/packages (A) and two single-locus programs (B). The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multi-locus programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. The median run times for the programs/packages across the scenarios are given in the table (C). The entries in a blue font  correspond to the lowest run-time for a scenario. FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
 FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
 }
\vspace{1cm}
%% plot formed via PowerPoint - ~/AM-Paper/Figure_power.pptx
%% script on mac
%% ~/AM-Paper/AM+/Plots_for_Paper/time_and_memory_plot_for_paper.R
\label{fig_time}
\centering
    \includegraphics[width=1.2\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
\end{figure}

@@

\begin{figure}
\caption{
\setstretch{1.1}{
Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level.  Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate of 5\% (Eagle$^{optimal}$).   The number of SNP-trait associations found are reported in the cells. }
}
\label{figmouse}
\begin{center}
%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
\includegraphics[width=12cm, height=12cm]{mouseresults.eps}
\end{center}

@@

\bibliographystyle{natbib}
%%\bibliographystyle{naturemag}
\bibliography{biblibrary}

>>>>

\documentclass{bioinfo}
\copyrightyear{2019} \pubyear{2019}

%%\usepackage{xr}
%%\renewcommand{\baselinestretch}{1.4} 
%%\usepackage{graphicx}
%%\usepackage{bm}
%%\usepackage{lineno}
%%\usepackage{natbib}
%%\usepackage{setspace}
%%\usepackage{authblk}
%%\usepackage{xcolor}

\access{Advance Access Publication Date: Day Month Year}
\appnotes{Original Paper}

\begin{document}
\firstpage{1}

\subtitle{Genetics and Population Analysis}

\title[Eagle: making multi-locus association mapping routine]{Eagle: multi-locus association mapping on a genome-wide scale made routine}
\author[Sample \textit{et~al}.]{Andrew W. George\,$^{\text{\sfb 1,}*}$, Arunas Verbyla\,$^{\text{\sfb 2}}$ and Joshua Bowden\,$^{\text{\sfb 3,}}$}
\address{$^{\text{\sf 1}}$Data61, CSIRO, Brisbane, 4102, Australia and \\
$^{\text{\sf 2}}$Data61, CSIRO, Atherton, 4883, Australia and \\
$^{\text{\sf 3}}$IM\&T, CSIRO, Brisbane, 4067, Australia.
}

\corresp{$^\ast$To whom correspondence should be addressed}

\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}

\editor{Associate Editor: Alfonso Valencia}

%%\linenumbers

\abstract{\textbf{Motivation:} We present Eagle, a new method for multi-locus association mapping. The motivation for developing Eagle was to make multi-locus association mapping "easy" and the method-of-choice. Eagle's strengths are that it a. is considerably more powerful than single-locus association mapping b. does not suffer from multiple testing issues c. gives results that are immediately interpretable and d. has a computational footprint comparable to single-locus association mapping. \\
\textbf{Results:} By conducting a large simulation study, we will show that Eagle finds true and avoids false SNP-trait associations better than competing single- and multi-locus methods. We also analyse data from a published mouse study. Eagle found over 50\% more validated findings than the state-of-the-art single-locus method.  \\
\textbf{Availability and Implementation:} Eagle has been implemented as an R package, with a browser-based Graphical User Interface (GUI) for users less familiar with R. It is freely available via the CRAN website  at https://cran.r-project.org. Videos, Quick Start guides, FAQs, and Demos are available via the Eagle website http://eagle.r-forge.r-project.org \\
\textbf{Contact:} \href{andrew.george@csiro.au}\\
\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics}
online.}

@@

\section{Introduction}

\citep{Boffelli03}

Over the past decade,  genome-wide association studies (GWASs) have changed considerably in both their analysis and design. Early studies
 followed a case-control design. Association mapping methods were no more complicated than contingency table tests or simple 
linear regression. These designs though had a tendency to yield spurious findings if there was unrecognised population stratification 
\citep{cardon2003population}. This prompted a shift towards family-based designs and score tests, such as the transmission/disequilibrium test (TDT)  and its variants \citep{spielman1996tdt}. Today, instead of by design, it is through statistical modelling that we account for the effects of population stratification \citep{price2010new}. This has meant that data can be collected from general populations, even if these populations are highly structured. Analysis via sophisticated association mapping methods, such as linear mixed model based approaches,  is now almost routine  \citep{yu2006unified,zhao2007arabidopsis}.

@@

%%\begin{figure}
%%\caption{
%%\setstretch{1.6}{
%%Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
%%FaST-LMM was run where all the SNP data are used 
%%to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
%%estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods.   }
%%}
%%\vspace{-3cm}
%%\label{figpowersingle}
%%\begin{center}
%%\includegraphics[width=10cm, height=15cm]{power2main.eps}
%%\end{center}

%%\end{figure}

\begin{figure}[!tpb]%figure2
\centerline{\includegraphics[width=8cm, height=10cm]{power2main.eps}}
\vspace{-2cm}
\caption{Power verse false positive rates for Eagle and the single-locus methods GEMMA and FaST-LMM. 
FaST-LMM was run where all the SNP data are used 
to estimate the relationship matrix (FaST-LMM$^{all}$)   and where genotype data from every five-hundredth SNP are used to 
estimate the relationship matrix (FaST-LMM$^{few}$). Eagle has substantially higher power than the single-locus methods. }
\label{figpowersingle}
\end{figure}

@@

\begin{figure}[!tpb]
 \centerline{\includegraphics[width=0.45\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}}
 \caption{
Median run times, in minutes, for the analysis of simulation study data from the six scenarios. Eagle is compared against five other multi-locus programs/packages (A) and two single-locus programs (B). The x- and y-axes are on a log scale for improved aesthetics. Eagle has the lowest run-times of the multi-locus programs/packages, sometimes by orders of magnitude. Eagle can even produce results faster than single-locus programs. The median run times for the programs/packages across the scenarios are given in the table (C). The entries in a blue font  correspond to the lowest run-time for a scenario. FaST-LMM$^{all}$ is where calculation of the similarity matrix is based on all the SNP data.  
 FaST-LMM$^{few}$ is where calculation  of the similarity matrix is based on a subset of the SNP data. }
\label{fig_time}
\end{figure}

@@

\begin{figure}[!tpb]
%% obtained from bracewell by running ResultsV2.R from /data/geo047/MWAM/Mouse/OutBred/RScripts/
\centerline{\includegraphics[width=8cm, height=8cm]{mouseresults.eps}}
\caption{
Genome-wide association mapping results from analyses of the mouse data for the single-locus method FaST-LMM and the 
multi-locus method Eagle. Genome-wide significance thresholds for FaST-LMM were calculated, via the Bonferroni correction 
(FaST-LMM$^{Bonf}$) and permutation (FaST-LMM$^{perm}$), at the 5\% significance level.  Eagle was run under three settings; its default setting (Eagle$^{default}$),  an alternate less conservative setting based on the number of SNPs and sample size (Eagle$^{alt}$), and where the model selection had been optimised for a false positive rate of 5\% (Eagle$^{optimal}$).   The number of SNP-trait associations found are reported in the cells. }
\label{figmouse}

@@

\begin{thebibliography}{}

\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03}
Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.

\bibitem[Ball, 2001]{ball2001bayesian}
Ball,R.D. (2001) Bayesian methods for quantitative trait loci mapping based on model
  selection: approximate analysis using the bayesian information criterion.
\textit{Genetics}, \textbf{159}, 1351--1364.

\bibitem[Boulesteix \textit{et~al.}, 2012]{boulesteix2012overview}
Boulesteix,A.-L.  \textit{et~al}. (2012)
\newblock Overview of random forest methodology and practical guidance with
  emphasis on computational biology and bioinformatics.
\newblock {\em Wiley Interdisciplinary Reviews: Data Mining and Knowledge
  Discovery\/}, {\bf 2}, 493--507.

\bibitem[Broman and Speed, 2002]{broman2002model}
Broman,K.W. and Speed,T.P. (2002)
\newblock A model selection approach for the identification of quantitative
  trait loci in experimental crosses.
\newblock {\em Journal of the Royal Statistical Society: Series B (Statistical
  Methodology)\/}, {\bf 64}, 641--656.

\bibitem[Browning, 2008]{browning2008presto}
Browning,B.L. (2008)
\newblock {PRESTO}: rapid calculation of order statistic distributions and
  multiple-testing adjusted {P}-values via permutation for one and two-stage
  genetic association studies.
\newblock {\em BMC Bioinformatics\/}, {\bf 9}, 309.

\bibitem[Cardon and Palmer, 2003]{cardon2003population}
Cardon,L.R. and Palmer,L.J. (2003)
\newblock Population stratification and spurious allelic association.
\newblock {\em The Lancet\/}, {\bf 361}, 598--604.

\bibitem[Chen and Chen(2008)]{chen2008extended}
Chen,J. and Chen,Z. (2008)
\newblock Extended {B}ayesian information criteria for model selection with
  large model spaces.
\newblock {\em Biometrika\/}, {\bf 95}, 759--771.

\bibitem[Cho \textit{et~al}., 2010]{cho2010joint}
Cho,S. \textit{et~al}. (2010) Joint identification of multiple genetic variants via elastic-net
  variable selection in a genome-wide association analysis. \textit{Annals of Human Genetics}, {\bf 74}(5), 416--428.

\bibitem[Consortium {\em et~al.}, 2010]{10002010map}
Consortium,G.P. {\em et~al.} (2010)
\newblock A map of human genome variation from population-scale sequencing.
\newblock {\em Nature\/}, {\bf 467}, 1061--1073.

\bibitem[de~Bakker {\em et~al.}, 2005]{de2005efficiency}
de~Bakker,P.I. \textit{et al.} (2005)
\newblock Efficiency and power in genetic association studies.
\newblock {\em Nature Genetics\/}, {\bf 37}, 1217--1223.

\bibitem[Doerge and Churchill, 1996]{doerge1996permutation}
Doerge,R.W. and Churchill,G.A. (1996)
\newblock Permutation tests for multiple loci affecting a quantitative
  character.
\newblock {\em Genetics\/}, {\bf 142}, 285--294.

\bibitem[Friedman {\em et~al.}, 2010]{Friedman2010glmnet}
Friedman,J. \textit{et al.} (2010)
\newblock Regularization {P}aths for {G}eneralized {L}inear {M}odels via
  {C}oordinate {D}escent.
\newblock {\em Journal of Statistical Software\/}, {\bf 33}, 1--22.

\bibitem[Gogel {\em et~al.}, 2018]{gogel2018comparison}
Gogel, B. \textit{et al.} (2018)
\newblock Comparison of a one-and two-stage mixed model analysis of
  {A}ustralia's {N}ational {V}ariety {T}rial {S}outhern {R}egion wheat data.
\newblock {\em Euphytica\/}, {\bf 214}, 44.

\bibitem[Golub and Van~Loan, 2012]{golub2012matrix}
Golub,G.H. and Van~Loan,C.F. (2012)
\newblock {\em Matrix {C}omputations\/}, volume~3.
\newblock JHU Press.

\bibitem[Gudbjartsson {\em et~al.}, 2015]{gudbjartsson2015large}
Gudbjartsson,D.F.  {\em et~al.} (2015)
\newblock Large-scale whole-genome sequencing of the {I}celandic population.
\newblock {\em Nature Genetics\/}, {\bf 47}, 435--444.

\bibitem[Kang {\em et~al.}, 2008]{kang2008efficient}
Kang,H.M. \textit{et al.} (2008)
\newblock Efficient control of population structure in model organism
  association mapping.
\newblock {\em Genetics\/}, {\bf 178}, 1709--1723.

\bibitem[Li and Ji, 2005]{li2005adjusting}
Li,J. and Ji,L. (2005)
\newblock Adjusting multiple testing in multilocus analyses using the
  eigenvalues of a correlation matrix.
\newblock {\em Heredity\/}, {\bf 95}, 221--227.

\bibitem[Lippert {\em et~al.}, 2011]{lippert2011fast}
Lippert,C. \textit{et al.} (2011)
\newblock {FaST} linear mixed models for genome-wide association studies.
\newblock {\em Nature Methods\/}, {\bf 8}, 833--835.

\bibitem[Liu {\em et~al.}, 2016]{liu2016iterative}
Liu, X. \textit{et al.} (2016)
\newblock Iterative usage of fixed and random effect models for powerful and
  efficient genome-wide association studies.
\newblock {\em PLoS Genetics\/}, {\bf 12}, e1005767.

\bibitem[Lockhart {\em et~al.}, 2014]{lockhart2014significance}
Lockhart,R. \textit{et al.}  (2014)
\newblock A significance test for the {L}asso.
\newblock {\em Annals of Statistics\/}, {\bf 42}, 2138--2139.

\bibitem[Long {\em et~al.}, 2017]{long2017whole}
LongT.  {\em et~al.} (2017)
\newblock Whole-genome sequencing identifies common-to-rare variants associated
  with human blood metabolites.
\newblock {\em Nature Genetics\/}, {\bf 49}, 568--578.

\bibitem[Meinshausen and B{\"u}hlmann, 2010]{meinshausen2010stability}
Meinshausen,N. and B{\"u}hlmann,P. (2010)
\newblock Stability {S}election.
\newblock {\em Journal of the Royal Statistical Society: Series B (Statistical
  Methodology)\/}, {\bf 72}, 417--473.

\bibitem[Nicod {\em et~al.}, 2016]{nicod2016genome}
Nicod,J.  {\em et~al.} (2016)
\newblock Genome-wide association of multiple complex traits in outbred mice by
  ultra-low-coverage sequencing.
\newblock {\em Nature Genetics\/},  {\bf 48}, 912--918.

\bibitem[Pahl and Schafer, 2010]{pahl2010permory}
Pahl,R. and Schafer,H. (2010)
\newblock {PERMORY}: an {LD}-exploiting permutation test algorithm for powerful
  genome-wide association testing.
\newblock {\em Bioinformatics\/}, {\bf 26}, 2093--2100.

\bibitem[Price {\em et~al.}, 2010]{price2010new}
Price,A.L. \textit{et al.} (2010)
\newblock New approaches to population stratification in genome-wide
  association studies.
\newblock {\em Nature Reviews Genetics\/}, {\bf 11}, 459--463.

\bibitem[Rakitsch {\em et~al.}, 2013]{rakitsch2013lasso}
Rakitsch,B. \textit{et al.} (2013)
\newblock A {L}asso multi-marker mixed model for association mapping with
  population structure correction.
\newblock {\em Bioinformatics\/}, {\bf 29}, 206--214.

\bibitem[Segura {\em et~al.}, 2012]{segura2012efficient}
Segura,V. \textit{et al.} (2012)
\newblock An efficient multi-locus mixed-model approach for genome-wide
  association studies in structured populations.
\newblock {\em Nature Genetics\/}, {\bf 44}, 825--830.

\bibitem[Shen {\em et~al.}, 2013]{shen2013novel}
Shen, X. \textit{et al.} (2013) 
\newblock A novel generalized ridge regression method for quantitative
  genetics.
\newblock {\em Genetics\/}, {\bf 193}, 1255--1268.

\bibitem[Spielman and Ewens, 1996]{spielman1996tdt}
Spielman,R.S. and Ewens,W.J. (1996)
\newblock The {TDT} and other family-based tests for linkage disequilibrium and
  association.
\newblock {\em American Journal of Human Genetics\/}, {\bf 59}, 983--989.

\bibitem[Storey and Tibshirani, 2003]{storey2003statistical}
Storey,J.D. and Tibshirani,R. (2003) 
\newblock Statistical significance for genomewide studies.
\newblock {\em Proceedings of the National Academy of Sciences\/}, {\bf
  100}, 9440--9445.

\bibitem[Szymczak {\em et~al.}, 2016]{szymczak2016r2vim}
Szymczak,S. \textit{et al.} (2016)
\newblock {r2VIM}: A new variable selection method for random forests in
  genome-wide association studies.
\newblock {\em BioData Mining\/}, {\bf 9}, 7.

\bibitem[Verbyla {\em et al.}, 2007]{verbyla2007analysis}
Verbyla,A.P. \textit{et al.} (2007)
\newblock The analysis of {QTL} by simultaneous use of the full linkage map.
\newblock {\em Theoretical and Applied Genetics\/}, {\bf 116}, 95--111.

\bibitem[Verbyla {\em et al.}, 2012]{verbyla2012rwgaim}
Verbyla,A.P. \textit{et al.} (2012)
\newblock {RWGAIM}: an efficient high-dimensional random whole genome average
  ({QTL}) interval mapping approach.
\newblock {\em Genetics Research\/}, {\bf 94}, 291--306.

\bibitem[Verbyla {\em et al.}, 2014]{verbyla2014whole}
Verbyla,A.P. \textit{et al.} (2014)
\newblock Whole-genome analysis of multienvironment or multitrait {QTL} in
  {MAGIC}.
\newblock {\em G3: Genes, Genomes, Genetics\/}, {\bf 4}, 1569--1584.

\bibitem[Yi {\em et~al.}, 2005]{yi2005bayesian}
Yi, N. \textit{et al.} (2005)
\newblock Bayesian model selection for genome-wide epistatic quantitative trait
  loci analysis.
\newblock {\em Genetics\/}, {\bf 170}, 1333--1344.

\bibitem[Yu {\em et~al.}, 2006]{yu2006unified}
Yu,J.  {\em et al.} (2006) 
\newblock A unified mixed-model method for association mapping that accounts
  for multiple levels of relatedness.
\newblock {\em Nature Genetics\/}, {\bf 38}, 203--208.

\bibitem[Zhao {\em et al.}, 2007]{zhao2007arabidopsis}
Zhao, K.  {\em et al.} (2007)
\newblock An {A}rabidopsis example of association mapping in structured
  samples.
\newblock {\em PLoS Genetics\/}, {\bf 3}, e4.

\bibitem[Zheng {\em et~al.}, 2011]{zheng2011comparison}
Zheng,J. \textit{et al.} (2011)
\newblock A comparison of approaches to account for uncertainty in analysis of
  imputed genotypes.
\newblock {\em Genetic Epidemiology\/}, {\bf 35}, 102--110.

\bibitem[Zhou and Stephens, 2012]{zhou2012genome}
Zhou,X. and Stephens,M. (2012)
\newblock Genome-wide efficient mixed-model analysis for association studies.
\newblock {\em Nature Genetics\/}, {\bf 44}, 821--824.

\end{thebibliography}
